---
title: "Diabetes-dementia updated simulation results"
author: "Andrew Mertens"
date: "2022-11-17"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, knitr.kable.NA = '')

source(here::here("0_config.R"))
source(paste0(here::here(),"/0_ltmle_Estimate_update.R"))
try(source(paste0(here::here(),"/simulation study/0_simulation_functions.R")))


load(paste0(here::here(),"/results/sim_performance_results_original.Rdata"))
load(paste0(here::here(),"/results/sim_performance_results_original_null.Rdata"))

library(knitr)
library(xtable)
library(kableExtra)

```


# To do:

  - Figure out why Qint and no Qint LASSO are equal and fix
  - Highlight DetQ lasso for both
  - Update the protective table caption on the



To change in organization
cgnage method to "Estimator" and put as first column
change estimator to "algorithm" and put as second column

NA for IPTW Qint and detq, but pick just detq (or not for qint)

Figure out why GLM Q-int and no Q-int is the same--- rerun

Add in different trunctation

-Fix lasso pre-screening labeling

# Scenario 1: Realistic simulation, null outcome

True RD: 0

```{r, echo=FALSE}

#Only use IC variance estimator
#res_null_diff <- res_null_diff %>% filter(variance_estimator=="ic")

#Set Qint to NA if IPTW
res_null_diff$Qint[res_null_diff$iptw=="IPTW"] <- "NA"

#Set DetQ to NA if IPTW and drop duplicates
res_null_diff$DetQ[res_null_diff$iptw=="IPTW"] <- "NA"



#colnames(res_null_diff)
#unique(res_null_diff$filenames)
res_null_diff_coverage <- res_null_diff %>% 
  distinct(estimator, iptw, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage, .keep_all = TRUE) %>%
  mutate(filenames=gsub("old_null_sim_res_","",filenames)) %>%
  #filter(variance_estimator=="ic") %>%
  #filter(!(filenames %in% c("noDetQ_Qint_tmle_T11")))  %>%
  # filter(filenames %in% c("old_null_sim_res_noDetQ_ic_T11","old_null_sim_res_noDetQ_Qint_ic_T11","old_null_sim_res_ic_glm_T11","old_null_sim_res_noDetQ_ic_glm_T11","old_null_sim_res_Qint_noDetQ_glm_tmle_T11")) %>%
  select(filenames, estimator, iptw, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage, ) %>% rename(oracle.coverage=o.coverage)

#drop duplicate filenames
res_null_diff_coverage <- res_null_diff_coverage %>% filter(!filenames %in% c("noDetQ_ic_T11","noDetQ_ic_glm_T11_iptw","noDetQ_tmle_T11_iptw","noDetQ_Qint_ic_T11_iptw","noDetQ_Qint_tmle_T11"))


res_null_diff_coverage$estimator <- "LASSO"
res_null_diff_coverage$estimator[grepl("rf",res_null_diff_coverage$filenames)] <- "Random Forest"
res_null_diff_coverage$estimator[grepl("glm",res_null_diff_coverage$filenames)] <- "GLM"
res_null_diff_coverage$estimator[grepl("ridge",res_null_diff_coverage$filenames)] <- "Ridge"
res_null_diff_coverage$estimator[grepl("EN",res_null_diff_coverage$filenames)] <- "Elastic Net"
res_null_diff_coverage$estimator[grepl("lasso_prescreen",res_null_diff_coverage$filenames)] <- "GLM, LASSO prescreen"
# res_null_diff_coverage$estimator[grepl("_DetQ",res_null_diff_coverage$filenames)] <- paste0(res_null_diff_coverage$estimator[grepl("_DetQ",res_null_diff_coverage$filenames)],", Det-Q")
# res_null_diff_coverage$estimator[grepl("_detQ",res_null_diff_coverage$filenames)] <- paste0(res_null_diff_coverage$estimator[grepl("_detQ",res_null_diff_coverage$filenames)],", Det-Q")


res_null_table <- res_null_diff_coverage  %>%
  filter(!is.na(variance)) %>% arrange(estimator,  Qint, DetQ, iptw, bias, variance) 


# identify index of rows to highlight
  row.i.1 <- which(res_null_table$filenames=="noDetQ_ic_T11")
  
res_null_table <- res_null_table %>% select(filenames, iptw, estimator, Qint, DetQ,bias,variance,
                                            #mse,
                                            bias_se_ratio,oracle.coverage) %>%
                  arrange(iptw, estimator,  Qint, DetQ,  bias, variance) %>%
                  rename(`Estimation method`=estimator,  `Q-int`=Qint, `Det. Q`=DetQ, Estimator=iptw, `Bias/SE`=bias_se_ratio,
                         Bias=bias, Variance=variance, `Oracle coverage`=oracle.coverage)

knitr::kable(res_null_table)

res_null_xtable <- res_null_table %>% 
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",  
    digits =5) %>%
  kableExtra::kable_styling(
      position = "left"#,
      # latex_options = c("striped", "repeat_header"),
      # stripe_color = "gray!15"
    ) %>%
  kable_styling()%>%
  row_spec(row.i.1-1, hline_after = T) %>%
  row_spec(row.i.1, bold=T,hline_after = T)

res_null_xtable

save_kable(res_null_xtable, file="C:/Users/andre/Documents/jici/diab-dementia-server-code/tables/null_sim_results_table.tex",float = FALSE)


```

  



# Scenario 2: Realistic simulation, protective effect of GLP1 on dementia 


True Risk Difference: -0.009683665


## Comparison of different estimators' performance

**Notes:**
  
  * Based on these results, we chose the LASSO estimator with Q-prediction and no deterministic Q function
  * Several of the estimators have comparable performance, but the chosen estimator performs best in both RR and RD estimation
  * Ridge regressions have lower MSE but not perfect 95% oracle coverage
  * Including the deterministic Q function marginally decreases bias/variance, so we should use in the bootstrap estimator
 

### Risk difference

```{r, echo=FALSE}

#colnames(res_v3_diff)
#unique(res_v3_diff$filenames)

res_v3_diff$DetQ[res_v3_diff$iptw=="IPTW"] <- "NA"
res_v3_diff$DetQ[res_v3_diff$iptw=="IPTW"] <- "NA"



res_v3_diff_coverage <- res_v3_diff %>% subset(., select = -c(bias_se_ratio_emp, power, censoring_in_data, mean_ci_width, simulated_data))
res_v3_diff_coverage <- res_v3_diff_coverage %>% filter(variance_estimator=="ic" | is.na(variance_estimator), !is.na(bias))

#unique(res_v3_diff_coverage$filenames)
res_v3_diff_coverage$estimator <- "LASSO"
res_v3_diff_coverage$estimator[grepl("rf",res_v3_diff_coverage$filenames)] <- "Random Forest"
res_v3_diff_coverage$estimator[grepl("glm",res_v3_diff_coverage$filenames)] <- "GLM"
res_v3_diff_coverage$estimator[grepl("ridge",res_v3_diff_coverage$filenames)] <- "Ridge"
res_v3_diff_coverage$estimator[grepl("EN",res_v3_diff_coverage$filenames)] <- "Elastic Net"
res_v3_diff_coverage$estimator[grepl("lasso_prescreen",res_v3_diff_coverage$filenames)] <- "GLM, LASSO prescreen"
# res_v3_diff_coverage$estimator[grepl("_DetQ",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_DetQ",res_v3_diff_coverage$filenames)],", Det-Q")
# res_v3_diff_coverage$estimator[grepl("_detQ",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_detQ",res_v3_diff_coverage$filenames)],", Det-Q")
#res_v3_diff_coverage$estimator[grepl("_iptw",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_iptw",res_v3_diff_coverage$filenames)],", IPTW")


#res_v3_diff_coverage$estimator[res_v3_diff_coverage$Qint=="Yes"] <- paste0(res_v3_diff_coverage$estimator[res_v3_diff_coverage$Qint=="Yes"],", Q-intercept")
res_v3_diff_coverage$estimator[grepl("1se",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("1se",res_v3_diff_coverage$filenames)],", Lambda: 1se")
res_v3_diff_coverage$estimator[grepl("AUC",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("AUC",res_v3_diff_coverage$filenames)],", AUC fit")

res_v3_diff_coverage <- res_v3_diff_coverage %>% 
  select(filenames, estimator, Qint, DetQ, iptw, bias, variance, mse, bias_se_ratio, o.coverage, filenames) %>% arrange(o.coverage)  %>% rename(oracle.coverage=o.coverage) %>% distinct()

#knitr::kable(res_v3_diff_coverage, digits =6)

# #remove duplicates /iptw repeats
# res_v3_diff <- res_v3_diff %>% filter(!filenames %in% c("sim_res_ridge_ic_v3_iptw","sim_res_ridge","sim_res_Qint_ic","sim_res_Qint_ic_v3_iptw",                                                    "sim_res_DetQ__ridge_ic_v3_iptw","sim_res_ridge_ic_v3","sim_res_DetQ_ic_v3_iptw","sim_res_ic"))



res_table <- res_v3_diff_coverage  %>%
  # #remove duplicates /iptw repeats
  filter(!(filenames %in% c("sim_res_noDetQ_Qint_tmle", "sim_res_Qint_ic_iptw","sim_res_Qint_AUC_iptw", "sim_res_ridge_AUC_iptw",
                            "sim_res_Qint_EN", "sim_res_Qint_1se_iptw",
                            "sim_res_DetQ_ic_v2","sim_res_DetQ_ic_v2_iptw", "sim_res_Qint_EN_iptw",
                            "sim_res_noDetQ_ic_v2_iptw","sim_res_noDetQ_Qint_ic_iptw"
                            )),
         estimator!="LASSO, Lambda: 1se, AUC fit", estimator!="Random Forest") %>%
  filter(!is.na(variance)) %>% arrange(estimator,  Qint, DetQ, iptw, bias, variance) 


# identify index of rows to highlight
  row.i.1 <- which(res_table$filenames=="sim_res_DetQ_ic_v3")
  
res_table <- res_table %>% select(filenames, 
                                  estimator,
                                  iptw, Qint, DetQ,
                                  bias,variance,
                                  #mse, #ask Maya if needed
                                  bias_se_ratio,oracle.coverage) %>%
                  arrange(iptw, estimator,  Qint, DetQ,  bias, variance) %>%
                  rename(`Estimation method`=estimator,  `Q-int`=Qint, `Det. Q`=DetQ, Estimator=iptw, `Bias/SE`=bias_se_ratio,
                         Bias=bias, Variance=variance, `Oracle coverage`=oracle.coverage)


knitr::kable(res_table)


#kable(res_table)

res_xtable <- res_table %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    digits =6) %>%
  kableExtra::kable_styling(
      position = "left"#,
      # latex_options = c("striped", "repeat_header"),
      # stripe_color = "gray!15"
    ) %>%
  kable_styling()%>%
  row_spec(row.i.1-1, hline_after = T) %>%
  row_spec(row.i.1, bold=T,hline_after = T)

res_xtable

save_kable(res_xtable, file="C:/Users/andre/Documents/jici/diab-dementia-server-code/tables/sim_results_table.tex",float = FALSE)


```




