---
title: "Diabetes-dementia updated simulation results"
author: "Andrew Mertens"
date: "2022-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("0_config.R"))
source(paste0(here::here(),"/0_ltmle_Estimate_update.R"))
try(source(paste0(here::here(),"/simulation study/0_simulation_functions.R")))
load(paste0(here::here(),"/simulation study/sim_performance_results/T4_null_sim_res.Rdata"))
load(paste0(here::here(),"/simulation study/sim_performance_results/T4_outcome_blind_sim_res.Rdata"))
load(paste0(here::here(),"/simulation study/sim_performance_results/T10_sim_res.Rdata"))
load(paste0(here::here(),"/results/simulation_results_old.Rdata"))

library(knitr)
```

## Null outcome- 1.5 year followup 

### Relative risk performance

```{r, echo=FALSE}
res <- perf_tab_RR_null_T4 %>% arrange(coverage)
knitr::kable(res, digits =3)


```

### Risk difference performance 

```{r, echo=FALSE}
res_diff <- perf_tab_diff_null_T4 %>% arrange(coverage)
knitr::kable(res_diff, digits =5)
```

### Notes

True RR=1, true RD=0
Bootstrap with clustered ID's in CV procedures

### Distribution of RR point estimates

```{r,echo=FALSE}
res_diff <- perf_tab_diff_null_T4 %>% arrange(coverage)
set.seed(123)
ggplot(d_T4_null, aes(y=analysis, x=estimate)) +
  geom_jitter(width=0, height=0.05, alpha=0.75) +
  coord_cartesian(xlim=c(0.4, 2)) +
  geom_vline(xintercept = 1) +
  geom_vline(aes(xintercept = true.RR), linetype="dashed") +
  scale_x_continuous(trans = "log10")

```



## Blind outcome- 1.5 year followup 

True RR:
True RD: 

### Relative risk performance

```{r, echo=FALSE}
res <- perf_tab_RR_outcome_blind_T4 %>% arrange(coverage)
knitr::kable(res, digits =3)


```

### Risk difference performance 

```{r, echo=FALSE}
res_diff <- perf_tab_diff_outcome_blind_T4 %>% arrange(coverage)
knitr::kable(res_diff, digits =5)
```

### Notes




## Blind outcome- 5 year followup 

True RR:
True RD: 

### Relative risk performance

```{r, echo=FALSE}
res <- perf_tab_RR_outcome_blind_T11 %>% arrange(coverage)
knitr::kable(res, digits =3)


```

### Risk difference performance 

```{r, echo=FALSE}
res_diff <- perf_tab_diff_outcome_blind_T11 %>% arrange(coverage)
knitr::kable(res_diff, digits =5)
```

### Notes




## Old results

True RR:
True RD: 


### Relative risk performance

```{r, echo=FALSE}
res <- perf_tab_RR_old %>% arrange(coverage)
knitr::kable(res, digits =3)


```

<!-- ### Risk difference performance  -->

<!-- ```{r, echo=FALSE} -->
<!-- res_diff <- perf_tab_diff_outcome_blind_T11 %>% arrange(coverage) -->
<!-- knitr::kable(res_diff, digits =5) -->
<!-- ``` -->


### Notes

