---
title: "Diabetes-dementia updated simulation results"
author: "Andrew Mertens"
date: "2022-11-17"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, knitr.kable.NA = '')

source(here::here("0_config.R"))
source(paste0(here::here(),"/0_ltmle_Estimate_update.R"))
try(source(paste0(here::here(),"/simulation study/0_simulation_functions.R")))


load(paste0(here::here(),"/results/sim_performance_results_original.Rdata"))
load(paste0(here::here(),"/results/sim_performance_results_original_null.Rdata"))

library(knitr)
library(xtable)
library(kableExtra)

```





# Scenario 1: Realistic simulation, null outcome

True RD: 0

```{r, echo=FALSE}

#colnames(res_null_diff)
#unique(res_null_diff$filenames)
res_null_diff_coverage <- res_null_diff %>% 
  distinct(estimator, iptw, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage, .keep_all = TRUE) %>%
  mutate(filenames=gsub("old_null_sim_res_","",filenames)) %>%
  #filter(variance_estimator=="ic") %>%
  filter(!(filenames %in% c("noDetQ_Qint_tmle_T11")))  %>%
  # filter(filenames %in% c("old_null_sim_res_noDetQ_ic_T11","old_null_sim_res_noDetQ_Qint_ic_T11","old_null_sim_res_ic_glm_T11","old_null_sim_res_noDetQ_ic_glm_T11","old_null_sim_res_Qint_noDetQ_glm_tmle_T11")) %>%
  select(filenames, estimator, iptw, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage, ) %>% rename(oracle.coverage=o.coverage)

res_null_diff_coverage$estimator <- "LASSO"
res_null_diff_coverage$estimator[grepl("rf",res_null_diff_coverage$filenames)] <- "Random Forest"
res_null_diff_coverage$estimator[grepl("glm",res_null_diff_coverage$filenames)] <- "GLM"
res_null_diff_coverage$estimator[grepl("ridge",res_null_diff_coverage$filenames)] <- "Ridge"
res_null_diff_coverage$estimator[grepl("EN",res_null_diff_coverage$filenames)] <- "Elastic Net"
res_null_diff_coverage$estimator[grepl("lasso_prescreen",res_null_diff_coverage$filenames)] <- "GLM, LASSO prescreen"
res_null_diff_coverage$estimator[grepl("_DetQ",res_null_diff_coverage$filenames)] <- paste0(res_null_diff_coverage$estimator[grepl("_DetQ",res_null_diff_coverage$filenames)],", Det-Q")
res_null_diff_coverage$estimator[grepl("_detQ",res_null_diff_coverage$filenames)] <- paste0(res_null_diff_coverage$estimator[grepl("_detQ",res_null_diff_coverage$filenames)],", Det-Q")


res_null_table <- res_null_diff_coverage  %>%
  filter(!is.na(variance)) %>% arrange(estimator,  Qint, DetQ, iptw, bias, variance) 


# identify index of rows to highlight
  row.i.1 <- which(res_null_table$filenames=="noDetQ_ic_T11")
  
res_null_table <- res_null_table %>% select(estimator, iptw, Qint, DetQ,bias,variance,mse,bias_se_ratio,oracle.coverage) %>%
                  rename(Estimator=estimator,  `Q-int`=Qint, `Det. Q`=DetQ, Method=iptw, `Bias/SE`=bias_se_ratio,
                         Bias=bias, Variance=variance, `Oracle coverage`=oracle.coverage)


res_null_xtable <- res_null_table %>% 
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",  
    digits =5) %>%
  kableExtra::kable_styling(
      position = "left"#,
      # latex_options = c("striped", "repeat_header"),
      # stripe_color = "gray!15"
    ) %>%
  kable_styling()%>%
  row_spec(row.i.1-1, hline_after = T) %>%
  row_spec(row.i.1, bold=T,hline_after = T)

res_null_xtable

save_kable(res_null_xtable, file="C:/Users/andre/Documents/jici/diab-dementia-server-code/tables/null_sim_results_table.tex",float = FALSE)


```

  
## Performance of difference variance estimators on null data

**Notes:**
  
  * Only showing LASSO estimator results-all estimator performances assessed in the realistic simulated data below.
  * Sanity-check on estimation performance on data with a known null association between GLP1 and dementia.
  * The IC variance estimator is anti-conservative and the TMLE variance estimator is conservative.
  * The bootstrap is anti-conservative but less so than the IC variance estimator.
  * The TMLE estimator is very conservative, with CI widths 8-10X that of the bootstrap.
  * The IPTW estimator is uniformly biased with overly-wide confidence intervals in all simulations (not shown).
  

```{r, echo=FALSE}

#colnames(res_null_diff)
#unique(res_null_diff$filenames)
res_null_diff$variance_estimator[is.na(res_null_diff$variance_estimator)] <- "bootstrap"
res_null_diff_coverage <- res_null_diff %>% 
  filter(estimator=="LASSO"& Qint=="No"& DetQ=="No" |is.na(bias)) %>%
  select(variance_estimator,coverage, mean_ci_width, filenames)

knitr::kable(res_null_diff_coverage, digits =5)

```





# Scenario 2: Realistic simulation, protective effect of GLP1 on dementia 


True Risk Difference: -0.009683665


## Comparison of different estimators' performance

**Notes:**
  
  * Based on these results, we chose the LASSO estimator with Q-prediction and no deterministic Q function
  * Several of the estimators have comparable performance, but the chosen estimator performs best in both RR and RD estimation
  * Ridge regressions have lower MSE but not perfect 95% oracle coverage
  * Including the deterministic Q function marginally decreases bias/variance, so we should use in the bootstrap estimator
 

### Risk difference

```{r, echo=FALSE}

#colnames(res_v3_diff)
#unique(res_v3_diff$filenames)
res_v3_diff_coverage <- res_v3_diff %>% subset(., select = -c(bias_se_ratio_emp, power, censoring_in_data, mean_ci_width, simulated_data))
res_v3_diff_coverage <- res_v3_diff_coverage %>% filter(variance_estimator=="ic" | is.na(variance_estimator), !is.na(bias))

#unique(res_v3_diff_coverage$filenames)
res_v3_diff_coverage$estimator <- "LASSO"
res_v3_diff_coverage$estimator[grepl("rf",res_v3_diff_coverage$filenames)] <- "Random Forest"
res_v3_diff_coverage$estimator[grepl("glm",res_v3_diff_coverage$filenames)] <- "GLM"
res_v3_diff_coverage$estimator[grepl("ridge",res_v3_diff_coverage$filenames)] <- "Ridge"
res_v3_diff_coverage$estimator[grepl("EN",res_v3_diff_coverage$filenames)] <- "Elastic Net"
res_v3_diff_coverage$estimator[grepl("lasso_prescreen",res_v3_diff_coverage$filenames)] <- "GLM, LASSO prescreen"
# res_v3_diff_coverage$estimator[grepl("_DetQ",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_DetQ",res_v3_diff_coverage$filenames)],", Det-Q")
# res_v3_diff_coverage$estimator[grepl("_detQ",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_detQ",res_v3_diff_coverage$filenames)],", Det-Q")
#res_v3_diff_coverage$estimator[grepl("_iptw",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("_iptw",res_v3_diff_coverage$filenames)],", IPTW")


#res_v3_diff_coverage$estimator[res_v3_diff_coverage$Qint=="Yes"] <- paste0(res_v3_diff_coverage$estimator[res_v3_diff_coverage$Qint=="Yes"],", Q-intercept")
res_v3_diff_coverage$estimator[grepl("1se",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("1se",res_v3_diff_coverage$filenames)],", Lambda: 1se")
res_v3_diff_coverage$estimator[grepl("AUC",res_v3_diff_coverage$filenames)] <- paste0(res_v3_diff_coverage$estimator[grepl("AUC",res_v3_diff_coverage$filenames)],", AUC fit")

res_v3_diff_coverage <- res_v3_diff_coverage %>% 
  select(filenames, estimator, Qint, DetQ, iptw, bias, variance, mse, bias_se_ratio, o.coverage, filenames) %>% arrange(o.coverage)  %>% rename(oracle.coverage=o.coverage) %>% distinct()

#knitr::kable(res_v3_diff_coverage, digits =6)


res_table <- res_v3_diff_coverage  %>%
  filter(!(filenames %in% c("sim_res_noDetQ_Qint_tmle",
                            "sim_res_DetQ_ic_v2","sim_res_DetQ_ic_v2_iptw"#, "sim_res_Qint_ic"
                            )),
         estimator!="LASSO, Lambda: 1se, AUC fit", estimator!="Random Forest") %>%
  filter(!is.na(variance)) %>% arrange(estimator,  Qint, DetQ, iptw, bias, variance) 


# identify index of rows to highlight
  row.i.1 <- which(res_table$filenames=="sim_res_DetQ_ic_v3")
  
res_table <- res_table %>% select(#filenames, 
                                  estimator,
                                  iptw, Qint, DetQ,
                                  bias,variance,mse,bias_se_ratio,oracle.coverage) %>%
                  rename(Estimator=estimator,  `Q-int`=Qint, `Det. Q`=DetQ, Method=iptw, `Bias/SE`=bias_se_ratio,
                         Bias=bias, Variance=variance, `Oracle coverage`=oracle.coverage)


#kable(res_table)

res_xtable <- res_table %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    digits =6) %>%
  kableExtra::kable_styling(
      position = "left"#,
      # latex_options = c("striped", "repeat_header"),
      # stripe_color = "gray!15"
    ) %>%
  kable_styling()%>%
  row_spec(row.i.1-1, hline_after = T) %>%
  row_spec(row.i.1, bold=T,hline_after = T)

res_xtable

save_kable(res_xtable, file="C:/Users/andre/Documents/jici/diab-dementia-server-code/tables/sim_results_table.tex",float = FALSE)


```







## Comparison of different variance estimators



**Notes:**
  
  * Showing LASSO estimator results with modeled Q (rather than intercept-only)
  * The IC variance estimator is anti-conservative and the TMLE variance estimator is conservative
  * The bootstrap is anti-conservative but less so than the IC variance estimator
  * The IPTW estimator is uniformly biased with overly-wide confidence intervals in all simulations (not shown)
  
### Risk difference coverage

```{r, echo=FALSE}

res_v3_diff_var <- res_v3_diff %>% subset(., select = -c(bias,variance,mse,bias_se_ratio, censoring_in_data, o.coverage, simulated_data))
res_v3_diff_var <- res_v3_diff_var %>% filter(Qint  =="No" , !grepl("iptw",filenames)|is.na(variance_estimator), !grepl("ridge",filenames), !grepl("AUC",filenames),
                                                          !grepl("EN",filenames), !grepl("1se",filenames), !grepl("glm",filenames), !grepl("_lasso_prescreen",filenames),
                                                          filenames!="im_res_DetQ_ic_v3",
                                                          filenames=="no DetQ_iptw-IPTW"|!grepl("iptw",filenames))


#Add iptw variance

resdf_noDetQ_ic <- readRDS(paste0(here::here(),"/data/sim_res_noDetQ_ic_v2.RDS"))
iptw_variance = mean((mean(resdf_noDetQ_ic$iptw.ate)-resdf_noDetQ_ic$iptw.ate)^2)
#      variance=mean((mean(log(estimate))-log(estimate))^2),



res_v3_diff_var$variance_estimator[grepl("_DetQ",res_v3_diff_var$filenames)] <- paste0(res_v3_diff_var$variance_estimator[grepl("_DetQ",res_v3_diff_var$filenames)],", Det-Q")
res_v3_diff_var <- res_v3_diff_var %>% mutate(variance_estimator=case_when(
  filenames=="DetQ" ~ "Bootstrap, Det Q function",
  filenames=="DetQ, 500 iter" ~ "Bootstrap, Det Q function, 500 iterations",
  filenames=="no DetQ" ~ "Bootstrap",
  filenames=="Ridge" ~ "Bootstrap-Ridge",
  filenames=="no DetQ_iptw-IPTW" ~ "Bootstrap- IPTW",
  variance_estimator==variance_estimator ~variance_estimator 
)) %>% distinct() #%>% filter(filenames!="DetQ, 500 iter")

res_v3_diff_var <- res_v3_diff_var %>% select(variance_estimator, coverage, mean_ci_width, power, bias_se_ratio_emp,filenames)



knitr::kable(res_v3_diff_var, digits =5)

```



