---
title: "Diabetes-dementia updated simulation results"
author: "Andrew Mertens"
date: "2022-10-19"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("0_config.R"))
source(paste0(here::here(),"/0_ltmle_Estimate_update.R"))
try(source(paste0(here::here(),"/simulation study/0_simulation_functions.R")))

# load(paste0(here::here(),"/simulation study/sim_performance_results/T4_null_sim_res.Rdata"))
# load(paste0(here::here(),"/simulation study/sim_performance_results/T4_outcome_blind_sim_res.Rdata"))
# load(paste0(here::here(),"/simulation study/sim_performance_results/T10_sim_res.Rdata"))

# load(paste0(here::here(),"/results/sim_performance_results.Rdata"))
# load(paste0(here::here(),"/results/simulation_results_old.Rdata"))
load(paste0(here::here(),"/results/sim_performance_results_simple.Rdata"))
load(paste0(here::here(),"/results/sim_performance_results_original.Rdata"))
load(paste0(here::here(),"/results/sim_performance_results_original_null.Rdata"))

library(knitr)
```


# Scenario 1: Simple simulation

True RR: 0.57

True RD: -0.3

**Notes:**

  * Both the GLM and LASSO estimators perform well, a sanity check on the ltmle LASSO implementation. 
  * All variance estimators perform well, with IC, TMLE, IPTW and 1000-iteration bootstrap perform similarly
  * Increasing bootstrap iterations from 200 to 1000 increased coverage closer to 95% in the RR estimation. 200 iterations was sufficient for RD.

### Relative risk performance

```{r, echo=FALSE}
#res <- sim_res_null_T4$perf_tab_RR %>% arrange(coverage) %>% filter(grepl("no_cens",filenames)) %>% subset(., select=-c(simulated_data, filenames, censoring_in_data, power))
knitr::kable(perf_tab_RR, digits =3)

```

### Risk difference performance

```{r, echo=FALSE}
#res_diff <- sim_res_null_T4$perf_tab_diff %>% arrange(coverage)  %>% arrange(coverage) %>% filter(grepl("no_cens",filenames)) %>% subset(., select=-c(simulated_data, filenames, censoring_in_data, power))
knitr::kable(perf_tab_RD, digits =5)
```



# Scenario 2: Realistic simulation, null outcome


True RR: 1

True RD: 0

**Notes:**

  * Outcome, death and censoring all jointly scrambled 
  * Oracle coverage is good for all estimators, but GLM estimators have more variance
  
### RD oracle coverage of different estimators
  
```{r, echo=FALSE}

#colnames(old_sim_res_null_diff)
#unique(old_sim_res_null_diff$filenames)
old_sim_res_null_diff_coverage <- old_sim_res_null_diff %>% #filter(variance_estimator=="ic") %>%
  filter(filenames %in% c("old_null_sim_res_noDetQ_ic_T11","old_null_sim_res_noDetQ_Qint_ic_T11","old_null_sim_res_ic_glm_T11","old_null_sim_res_noDetQ_ic_glm_T11","old_null_sim_res_Qint_noDetQ_glm_tmle_T11")) %>%
  select(estimator, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage)

knitr::kable(old_sim_res_null_diff_coverage, digits =5)

```
  
### RR oracle coverage of different estimators
  
```{r, echo=FALSE}

old_sim_res_null_RR_coverage <- old_sim_res_null_RR %>% #filter(variance_estimator=="ic") %>%
  filter(filenames %in% c("old_null_sim_res_noDetQ_ic_T11","old_null_sim_res_noDetQ_Qint_ic_T11","old_null_sim_res_ic_glm_T11","old_null_sim_res_noDetQ_ic_glm_T11","old_null_sim_res_Qint_noDetQ_glm_tmle_T11")) %>%
  select(estimator, Qint, DetQ,bias,variance,mse,bias_se_ratio,o.coverage)

knitr::kable(old_sim_res_null_RR_coverage, digits = 3)

```
  
  
## Performance of difference variance estimators on null data

**Notes:**
  
  * Only showing LASSO estimator results-all estimator performances assessed in the realistic simulated data below
  * Sanity-check on estimation performance on data with a known null association between GLP1 and dementia
  * The IC variance estimator is anti-conservative and the TMLE variance estimator is conservative
  * The bootstrap is anti-conservative but less so than the IC variance estimator
  * The IPTW estimator is uniformly biased with overly-wide confidence intervals in all simulations (not shown)
  
### Risk difference performance

```{r, echo=FALSE}

#colnames(old_sim_res_null_diff)
#unique(old_sim_res_null_diff$filenames)
old_sim_res_null_diff$variance_estimator[is.na(old_sim_res_null_diff$variance_estimator)] <- "bootstrap"
old_sim_res_null_diff_coverage <- old_sim_res_null_diff %>% 
  filter(estimator=="LASSO"& Qint=="No"& DetQ=="No" |is.na(bias)) %>%
  select(variance_estimator,coverage, mean_ci_width)

knitr::kable(old_sim_res_null_diff_coverage, digits =5)

```


### Relative risk performance

```{r, echo=FALSE}

old_sim_res_null_RR$variance_estimator[is.na(old_sim_res_null_RR$variance_estimator)] <- "bootstrap"
old_sim_res_null_RR_coverage <- old_sim_res_null_RR %>% filter(estimator=="LASSO"& Qint=="No"& DetQ=="No" |is.na(bias)) %>%
  filter(!grepl("iptw",filenames)) %>%
  select(variance_estimator,coverage, mean_ci_width)

knitr::kable(old_sim_res_null_RR_coverage, digits =5)

```

Note CI width on the log scale.




# Scenario 3: Realistic simulation, protective effect of GLP1 on dementia 


True RD: -0.009683665

True RR: 0.5148661

## Comparison of different estimators' performance

**Notes:**
  
  * Based on these results, we chose the LASSO estimator with Q-prediction and no deterministic Q function
  * Several of the estimators have comparable performance, but the chosen estimator performs best in both RR and RD estimation
  * Ridge regressions have lower MSE but not perfect 95% oracle coverage
  * Including the deterministic Q function marginally decreases bias/variance, so we should use in the bootstrap estimator
 

### Risk difference

```{r, echo=FALSE}

#colnames(old_sim_res_v3_diff)
#unique(old_sim_res_v3_diff$filenames)
old_sim_res_v3_diff_coverage <- old_sim_res_v3_diff %>% subset(., select = -c(bias_se_ratio_emp, power, censoring_in_data, mean_ci_width, simulated_data))
old_sim_res_v3_diff_coverage <- old_sim_res_v3_diff_coverage %>% filter(variance_estimator=="ic" | is.na(variance_estimator), !is.na(bias))

#unique(old_sim_res_v3_diff_coverage$filenames)
old_sim_res_v3_diff_coverage$estimator <- "LASSO"
old_sim_res_v3_diff_coverage$estimator[grepl("glm",old_sim_res_v3_diff_coverage$filenames)] <- "GLM"
old_sim_res_v3_diff_coverage$estimator[grepl("ridge",old_sim_res_v3_diff_coverage$filenames)] <- "Ridge"
old_sim_res_v3_diff_coverage$estimator[grepl("EN",old_sim_res_v3_diff_coverage$filenames)] <- "Elastic Net"
old_sim_res_v3_diff_coverage$estimator[grepl("lasso_prescreen",old_sim_res_v3_diff_coverage$filenames)] <- "GLM, LASSO prescreen"
old_sim_res_v3_diff_coverage$estimator[grepl("_DetQ",old_sim_res_v3_diff_coverage$filenames)] <- paste0(old_sim_res_v3_diff_coverage$estimator[grepl("_DetQ",old_sim_res_v3_diff_coverage$filenames)],", Det-Q")
old_sim_res_v3_diff_coverage$estimator[old_sim_res_v3_diff_coverage$Qint=="Yes"] <- paste0(old_sim_res_v3_diff_coverage$estimator[old_sim_res_v3_diff_coverage$Qint=="Yes"],", Q-intercept")
old_sim_res_v3_diff_coverage$estimator[grepl("1se",old_sim_res_v3_diff_coverage$filenames)] <- paste0(old_sim_res_v3_diff_coverage$estimator[grepl("1se",old_sim_res_v3_diff_coverage$filenames)],", Lambda: 1se")
old_sim_res_v3_diff_coverage$estimator[grepl("AUC",old_sim_res_v3_diff_coverage$filenames)] <- paste0(old_sim_res_v3_diff_coverage$estimator[grepl("AUC",old_sim_res_v3_diff_coverage$filenames)],", AUC fit")

old_sim_res_v3_diff_coverage <- old_sim_res_v3_diff_coverage %>% 
  select(estimator, bias, variance, mse, o.coverage) %>% arrange(o.coverage)

knitr::kable(old_sim_res_v3_diff_coverage, digits =6)

```



### Relative Risk 

```{r, echo=FALSE}

#colnames(old_sim_res_v3_diff)
#unique(old_sim_res_v3_diff$filenames)

old_sim_res_v3_RR_coverage <- old_sim_res_v3_RR %>% subset(., select = -c(bias_se_ratio_emp, coverage, power, censoring_in_data, mean_ci_width, simulated_data))
old_sim_res_v3_RR_coverage <- old_sim_res_v3_RR_coverage %>% filter(variance_estimator=="ic" | is.na(variance_estimator), !is.na(bias), !grepl("iptw",filenames))

#unique(old_sim_res_v3_RR_coverage$filenames)
old_sim_res_v3_RR_coverage$estimator <- "LASSO"
old_sim_res_v3_RR_coverage$estimator[grepl("glm",old_sim_res_v3_RR_coverage$filenames)] <- "GLM"
old_sim_res_v3_RR_coverage$estimator[grepl("ridge",old_sim_res_v3_RR_coverage$filenames)] <- "Ridge"
old_sim_res_v3_RR_coverage$estimator[grepl("EN",old_sim_res_v3_RR_coverage$filenames)] <- "Elastic Net"
old_sim_res_v3_RR_coverage$estimator[grepl("lasso_prescreen",old_sim_res_v3_RR_coverage$filenames)] <- "GLM, LASSO prescreen"
old_sim_res_v3_RR_coverage$estimator[grepl("_DetQ",old_sim_res_v3_RR_coverage$filenames)] <- paste0(old_sim_res_v3_RR_coverage$estimator[grepl("_DetQ",old_sim_res_v3_RR_coverage$filenames)],", Det-Q")
old_sim_res_v3_RR_coverage$estimator[old_sim_res_v3_RR_coverage$Qint=="Yes"] <- paste0(old_sim_res_v3_RR_coverage$estimator[old_sim_res_v3_RR_coverage$Qint=="Yes"],", Q-intercept")
old_sim_res_v3_RR_coverage$estimator[grepl("1se",old_sim_res_v3_RR_coverage$filenames)] <- paste0(old_sim_res_v3_RR_coverage$estimator[grepl("1se",old_sim_res_v3_RR_coverage$filenames)],", Lambda: 1se")
old_sim_res_v3_RR_coverage$estimator[grepl("AUC",old_sim_res_v3_RR_coverage$filenames)] <- paste0(old_sim_res_v3_RR_coverage$estimator[grepl("AUC",old_sim_res_v3_RR_coverage$filenames)],", AUC fit")

old_sim_res_v3_RR_coverage <- old_sim_res_v3_RR_coverage %>% 
  select(estimator, bias, variance, mse, o.coverage) %>% arrange(o.coverage)

knitr::kable(old_sim_res_v3_RR_coverage, digits =3)

```
  


## Comparison of different variance estimators



**Notes:**
  
  * Showing LASSO estimator results with modeled Q (rather than intercept-only)
  * The IC variance estimator is anti-conservative and the TMLE variance estimator is conservative
  * The bootstrap is anti-conservative but less so than the IC variance estimator
  * The IPTW estimator is uniformly biased with overly-wide confidence intervals in all simulations (not shown)
  
### Risk difference coverage

```{r, echo=FALSE}

old_sim_res_v3_diff_var <- old_sim_res_v3_diff %>% subset(., select = -c(bias,variance,mse,bias_se_ratio, censoring_in_data, o.coverage, simulated_data))
old_sim_res_v3_diff_var <- old_sim_res_v3_diff_var %>% filter(Qint  =="No" , !grepl("iptw",filenames)|is.na(variance_estimator), !grepl("ridge",filenames), !grepl("AUC",filenames),
                                                          !grepl("EN",filenames), !grepl("1se",filenames), !grepl("glm",filenames), !grepl("_lasso_prescreen",filenames),
                                                          filenames!="im_res_DetQ_ic_v3",
                                                          filenames=="no DetQ_iptw-IPTW"|!grepl("iptw",filenames))

old_sim_res_v3_diff_var$variance_estimator[grepl("_DetQ",old_sim_res_v3_diff_var$filenames)] <- paste0(old_sim_res_v3_diff_var$variance_estimator[grepl("_DetQ",old_sim_res_v3_diff_var$filenames)],", Det-Q")
old_sim_res_v3_diff_var <- old_sim_res_v3_diff_var %>% mutate(variance_estimator=case_when(
  filenames=="DetQ" ~ "Bootstrap, Det Q function",
  filenames=="DetQ, 500 iter" ~ "Bootstrap, Det Q function, 500 iterations",
  filenames=="no DetQ" ~ "Bootstrap",
  filenames=="no DetQ_iptw-IPTW" ~ "Bootstrap- IPTW",
  variance_estimator==variance_estimator ~variance_estimator 
))

old_sim_res_v3_diff_var <- old_sim_res_v3_diff_var %>% select(variance_estimator, coverage, mean_ci_width, power, bias_se_ratio_emp)



knitr::kable(old_sim_res_v3_diff_var, digits =5)

```


  
### Relative risk coverage

```{r, echo=FALSE}

old_sim_res_v3_RR_var <- old_sim_res_v3_RR %>% subset(., select = -c(bias,variance,mse,bias_se_ratio, censoring_in_data, o.coverage, simulated_data))
old_sim_res_v3_RR_var <- old_sim_res_v3_RR_var %>% filter(Qint  =="No" , !grepl("iptw",filenames)|is.na(variance_estimator), !grepl("ridge",filenames), !grepl("AUC",filenames),
                                                          !grepl("EN",filenames), !grepl("1se",filenames), !grepl("glm",filenames), !grepl("_lasso_prescreen",filenames),
                                                          filenames!="im_res_DetQ_ic_v3",
                                                          filenames=="no DetQ_iptw-IPTW"|!grepl("iptw",filenames))

old_sim_res_v3_RR_var$variance_estimator[grepl("_DetQ",old_sim_res_v3_RR_var$filenames)] <- paste0(old_sim_res_v3_RR_var$variance_estimator[grepl("_DetQ",old_sim_res_v3_RR_var$filenames)],", Det-Q")
old_sim_res_v3_RR_var <- old_sim_res_v3_RR_var %>% mutate(variance_estimator=case_when(
  filenames=="DetQ" ~ "Bootstrap, Det Q function",
  filenames=="DetQ, 500 iter" ~ "Bootstrap, Det Q function, 500 iterations",
  filenames=="no DetQ" ~ "Bootstrap",
  filenames=="no DetQ_iptw-IPTW" ~ "Bootstrap- IPTW",
  variance_estimator==variance_estimator ~variance_estimator 
))

old_sim_res_v3_RR_var <- old_sim_res_v3_RR_var %>% select(variance_estimator, coverage, mean_ci_width, power, bias_se_ratio_emp)



knitr::kable(old_sim_res_v3_RR_var, digits =3)

```
